
---
title: "Models of Neighborhood Change in Raleigh-Durham-Chapel Hill"
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    source: embed
    smart: false
runtime: shiny
---

    

```{r global, echo=F}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )
library( rgdal )

```


```{r, include=FALSE}

# DATA STEPS 

# load dorling cartogram from github
# map already contains census data and groups from clustering 



# from local file path
rdch <- geojson_read( "/Users/asiawoods/Documents/GitHub/CPP529/rdch_dorling.geojson", what="sp" )





# reproject the map 
rdch2 <- spTransform( rdch, CRS("+init=epsg:3395") )

# convert the sp map format to 
# an sf (simple features) format:
# ggmap requires the sf format
rdch.sf <- st_as_sf( rdch2 )

# separate out the data frame from the map
d <- as.data.frame( rdch.sf )

plot(rdch)

```



```{r}
URL1 <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-2000.rds"
d1 <- readRDS( gzcon( url( URL1 ) ) )

URL2 <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-2010.rds"
d2 <- readRDS( gzcon( url( URL2 ) ) )

URLmd <- "https://github.com/DS4PS/cpp-529-fall-2020/raw/main/LABS/data/rodeo/LTDB-META-DATA.rds"
md <- readRDS( gzcon( url( URLmd ) ) )

d1 <- select( d1, - year )
d2 <- select( d2, - year )

d3 <- merge( d1, d2, by="tractid" )
d3 <- merge( d3, md, by="tractid" )

#Filter Urban Areas
d3 <- filter( d3, urban == "urban" )


# STANDARDIZE GEO IDs

# note the current geoid format for the LTDB census data: 
# FIPS-STATE-COUNTY-TRACT:  fips-01-001-020100  

x <- d3$tractid 

# remove non-numeric strings 
x <- gsub( "fips", "", x )
x <- gsub( "-", "", x )

# drop leading zeros 
x <- as.numeric( x )

# remember to add the variable back to the census dataset
d3$tractid2 <- x 

rdch2 <- merge( rdch2, d3, by.x="GEOID", by.y="tractid", all.x=T )
```

```{r}
d3 <- select( d3, tractid, 
             mhmval00, mhmval12, 
             hinc00, 
             hu00, vac00, own00, rent00, h30old00,
             empclf00, clf00, unemp00, prof00,  
             dpov00, npov00,
             ag25up00, hs00, col00, 
             pop00.x, nhwht00, nhblk00, hisp00, asian00,
             cbsa, cbsaname )

 
d3 <- 
  d3 %>%
  mutate( # percent white in 2000
          p.white = 100 * nhwht00 / pop00.x,
          # percent black in 2000
          p.black = 100 * nhblk00 / pop00.x,
          # percent hispanic in 2000
          p.hisp = 100 * hisp00 / pop00.x, 
          # percent asian in 2000
          p.asian = 100 * asian00 / pop00.x,
          # percent high school grads by age 25 in 2000 
          p.hs = 100 * (hs00+col00) / ag25up00,
          # percent pop with college degree in 2000
          p.col = 100 * col00 / ag25up00,
          # percent employed in professional fields in 2000
          p.prof = 100 * prof00 / empclf00,
          # percent unemployment  in 2000
          p.unemp = 100 * unemp00 / clf00,
          # percent of housing lots in tract that are vacant in 2000
          p.vacant = 100 * vac00 / hu00,
          # dollar change in median home value 2000 to 2010 
          pov.rate = 100 * npov00 / dpov00 )



# adjust 2000 home values for inflation 
mhv.2000 <- d3$mhmval00 * 1.28855  
mhv.2010 <- d3$mhmval12


# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $1,000 are outliers
mhv.2000[ mhv.2000 < 1000 ] <- NA

# change in MHV in dollars
mhv.change <- mhv.2010 - mhv.2000

# drop cases where value decrease by more than -$100,000 
mhv.change[ mhv.change < -100000 ] <- NA


# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.2000 )

# drop cases where growth exceeds 200% 
mhv.growth[ mhv.growth > 200 ] <- NA 


d3$mhv.2000 <- mhv.2000
d3$mhv.2010 <- mhv.2010
d3$mhv.change <- mhv.change
d3$mhv.growth <- mhv.growth

head(d3)
```


Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
these.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")


# replace these with descriptive labels 
prop.names <- c("Percent white, non-Hispanic", 
"Percent black, non-Hispanic", "Percent Hispanic", "Percent Native American race", 
"Percent foreign born", "Percent speaking other language at home, age 5 plus", 
"Percent with high school degree or less", "Percent with 4-year college degree or more", 
"Percent unemployed", "Percent female labor force participation", 
"Percent professional employees", "Percent manufacturing employees", 
"Percent veteran", "Percent self-employed", "Median HH income, total", 
"Per capita income", "Percent in poverty, total", "Percent owner-occupied units", 
"Percent vacant units", "Percent multi-family units", "Median rent", 
"Median home value", "Percent structures more than 30 years old", 
"Percent HH in neighborhood 10 years or less", "Percent 17 and under, total", 
"Percent 60 and older, total", "Percent 75 and older, total", 
"Percent currently married, not separated", "Percent widowed, divorced and separated", 
"Percent female-headed families with children")

x <- prop.names
names(x) <- these.variables
#x[ input$demographics ]

# from the data dictionary 
temp.names <- paste0( "Variable ", prop.names )

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              # choices = these.variables, 
              choiceNames=temp.names,
              choiceValues=these.variables,
              selected="pnhwht12")



# Adding interpretable variable names
# from the data dictionary:
# add a name attribute for each variable
# 
# value <- c(1,2,3)
# dd.name <- c("one","two","three")
# 
# x <- dd.name
# names(x) <- value
#
# dd names and values linked
# names( x[2] )
#
# can now get the label using the value
# using the name attributes 
# x[ "two" ]
#
# to add labels to the maps
# use the radio button value 
# to get the data dictionary label: 
#
# x[ input$demographics ]
```



Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             rdch.sf <- 
             rdch.sf %>% 
             mutate( q = ntile( get(input$demographics), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -12519146, xmax = -12421368 ) + 
    ylim( ymin = 3899074, ymax = 3965924 )

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 
bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                  ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395"))


# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

rdch2$cluster[ rdch2$cluster == "1" ] <- "Non-Hispanic White, married, self-employed professionals who are home-owners with at least a Bachelor’s Degree"
rdch2$cluster[ rdch2$cluster == "2" ] <- "Multilingual, unemployed Native Americans with at least high school degrees that live in poverty
"
rdch2$cluster[ rdch2$cluster == "3" ] <- "Single-parent Native Americans that are female-headed, multi-family households and live in poverty"
rdch2$cluster[ rdch2$cluster == "4" ] <- "Elderly Non-Hispanic White Veterans"
rdch2$cluster[ rdch2$cluster == "5" ] <- "Non-Hispanic Black minors that live in neighborhoods created in the last ten years"
rdch2$cluster[ rdch2$cluster == "6" ] <- "Married  professionals with at least a Bachelor’s Degree"
rdch2$cluster[ rdch2$cluster == "7" ] <- "Multilingual, foreign Hispanic persons who live in poverty and have at least a high school diploma"
rdch2$cluster[ rdch2$cluster == "8" ] <- "Non-married persons who live in 30+ year structures and work in a primarily female labor force
"



renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( rdch2, bbox=bb ) + 
  tm_polygons( col="cluster", palette="Accent",
               title="Community Types" ) 

})
```




NH Change 2000-2010 
===================================== 


Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.2000","mhv.2010","mhv.change","mhv.growth")

x <- button.values
names(x) <- button.labels

temp.button <- paste0("Variable ", button.values)
  
radioButtons( inputId="home.value", 
              label = h3("Home Values"),
             # choices = button.values, 
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.2000")



```









Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             rdch.sf <- 
             rdch.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -12519146, xmax = -12421368 ) + 
    ylim( ymin = 3899074, ymax = 3965924 )

})

```


### Variable Distribution 



```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] })

x <- get_variable_x() %>% unlist() %>% as.numeric()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```  







Drivers of Change   
===================================== 

Change variable names

Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.2000","mhv.2010","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.change")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

covariate.labels <- c("Percent white, non-Hispanic", 
"Percent black, non-Hispanic", "Percent Hispanic", "Percent Native American race", 
"Percent foreign born", "Percent speaking other language at home, age 5 plus", 
"Percent with high school degree or less", "Percent with 4-year college degree or more", 
"Percent unemployed", "Percent female labor force participation", 
"Percent professional employees", "Percent manufacturing employees", 
"Percent veteran", "Percent self-employed", "Median HH income, total", 
"Per capita income", "Percent in poverty, total", "Percent owner-occupied units", 
"Percent vacant units", "Percent multi-family units", "Median rent", 
"Median home value", "Percent structures more than 30 years old", 
"Percent HH in neighborhood 10 years or less", "Percent 17 and under, total", 
"Percent 60 and older, total", "Percent 75 and older, total", 
"Percent currently married, not separated", "Percent widowed, divorced and separated", 
"Percent female-headed families with children")

checkboxGroupInput( inputId="covariates", 
              label = h3("Select Variables for Your Model"),
              choices = covariates,
              choiceNames=covariate.labels,
              choiceValues=covariates,
              selected=c("pnhwht12","pprof12","pvac12") )

```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

# RUNNING A REGRESSION WITH USER INPUTS
#
# create a formula object
# by constructing the formula from user selections 
# as a string then casting as a formula object 
# x.s <- c("x1","x2","x3" )
# formula.text <- paste( "y", " ~ ", paste0( x.s, collapse=" + ") )
# formula.object <- as.formula( formula.text )
#
# lm( formula.object, data=d )
#
# make sure all variables are in d


  
# check boxes return vectors 
get_covariates <-  reactive({ input$covariates })


renderUI({
  
covariates <- get_covariates() 

formula.text <- paste0( "mhmval12", "mhv.change ~ dv ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )


})

# HTML( reg.table )
```



### Correlation Plots 

```{r}
pairs( iris )
```





<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>